\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-08 paper/draft.tex
Resolve comments

% To have any hope of parallelizing scripts, \sys first needs to understand the parallelization characteristics of the commands they contain.
\sys aims at parallelizing data-parallel commands, \ie ones that can process their input in parallel, encoding their characteristics by assigning them to
\emph{parallelizability classes}.
% \kk{eurosys: I have two small concerns about this sentence but I don't know how to fix them. First, there is a redundancy "parallelizing data-parallel". Second and most important, I would like this sentence to say that \sys exploits latent data parallelism in commands, but not that it parallelizes data parallel commands. The reason is that I would like to make sure that no reviewer thinks that all we do is parallelizing each command on its own (as some OSDI reviewer thought).}
%% This information is encoded in a command's \emph{parallelizability class}.
% Intuitively, a parallelizability 
A class represents the level of synchronization required by copies of a command executing in parallel, 
  capturing only characteristics that are important for its parallel execution. % , rather than the command's full observable behavior.
% \kk{eurosys: This sentence doesn't add much, especially given that we haven't talked about synchronization.}
\sys leans towards having a few coarse classes rather than many detailed ones---among other reasons, to simplify their understanding and use by command developers.
% Make the connections that this is like a type, that doesn't capture everything, but is good-enough

@@

% As parallelizability classes form a hierarchy from most parallelizable to least parallelizable~\sx{cmd}, a command is classified by the class of its least parallelizable flag.
% For example, if a custom \ttt{trace-sort} command is invoked with flags \ttt{r}, \ttt{n}, and \ttt{k2} that are in \pur and a flag \ttt{d} in \sid that writes debugging output to a file, it ends up in \sid.
% \kk{eurosys: I removed this paragraph because it is only a simplistic description of how flags are handled.}

@@

\heading{Example Annotations}
The parallelizability properties of each command are specified in an annotation record that contains several components, each addressing one of the key concerns.
On the top level, each annotation record contains a set of clauses, each one identified by a predicate on the command arguments (Concern (iii)). 
Each clause contains an assignment for the command, indicating its class (Concern (i)), and the sequence of its inputs and outputs (Concern (ii)).
%  (a JSON object) that
% Annotations treat flag arguments differently from file arguments, by checking if they start with a \ttt{-}.
% \footnote{
%   Note that the single character \ttt{-} is not an option as it is most often used to designate that a command reads from its standard input.
%   \nv{
%     There are options of of the form --f "x; y", and BSD options do not take \ttt{-} (like \ttt{ps aux}).
%     I am also not sure how important this last sentence is, as it did not answer a question I had---I expected arguments to be anything: if soneome wants to put ``pizza'' the DSL should not disallow it.
%   }
% }
% The complete annotation grammar is shown in Appendix~\ref{alg}. \str
% \kk{EuroSys doesn't allow appendix. I think we should just remove the grammar.}

@@

\heading{Translation Pass}
%
%% \kk{It might be a good idea to change this subsection based on the new
%%   as-late-as-possible architecture where everything is wrapped and
%%   translated at the final moment. Ideally, this would become the main
%%   point of this subsection, as the translation algorithm is kind of
%%   redundant when having the distributable regions too.}
%
\sys's front-end performs a depth-first search on the AST of
the given shell program.  During this pass, it extends the
dataflow regions bottom-up, translating their
independent components to DFG nodes until a barrier
construct is reached. All AST subtrees not translatable to DFGs are kept as they are. The output of the translation pass is the
original AST where dataflow regions have been replaced with
DFGs. 
% and calls to \sys's runtime primitives~\sx{optimizer}.
% \kk{eurosys: I removed this since it is slightly inaccurate and readers don't know what the primitives are at that point.}

@@

\heading{Highlights}
\sys accelerates ($1.92-61.1\times$) all scripts ($>50$, with $>200$ commands).
\kk{Add speedup separated: (i) transformations, (ii) runtime primitives.}
% From the rest, only two see a slowdown, only because \sys's setup ($\sim1s$) is higher than their runtime ($\sim0.4s$).
\sys's optimized primitives offer benefits even against manually parallelized scripts. % (such as GNU sort, and GNU parallel).
% \sys's compilation pass is negligible (avg. $<1s$). %  for non trivial program execution.
In two large cases, tasks such as data download, extraction, and preprocessing---often implemented using shell scripts and outside the focus of specialized parallelization frameworks---take comparable time to the execution of the main computation, and can be significantly accelerated by \sys.
% In terms of manual effort,

@@

\heading{Results}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents \sys's speedup as a function of parallelism (2--64$\times$).
% (Bi-grams-opt is shown as the \ttt{opt} line in Fig.~\ref{fig:microbenchmarks}'s Bi-grams.)
Average speedups of the optimized \sys, \ie \ttt{eager} and \ttt{split} enabled, for \{2, 4, 8, 16, 32, 64\}$\times$-parallelism are \todo{1.97, 3.5, 5.78, 8.83, 10.96, and 13.47}, respectively.
% and the geometric means are 1.92, 3.39, 5.48, 7.94, 8.86, 8.51.}
% \kk{For the one-liners where split doesn't provide any benefit (or overhead) we don't show the PaSh w/o split line.}
%
% we experiment with a $200\times$ level of parallelism to see potential overheads
Tab.~\ref{tab:eval} shows that \sys's compilation time is negligible. % ($<5s$)
It also contains filter and bigrams-opt (not in Fig.~\ref{fig:microbenchmarks}). The first (filter) is used to evaluate \sys's impact on IO-intensive scripts (\sys achieves a 1.5-2.5$\times$ speedup). The second (bigrams-opt) is an optimized version of bigrams where we grouped 5 commands into one \tpur command thas has a more efficient $map$ and $aggregate$ implementation. \sys achieves up to x30 speedup for this script (significantly higher than what is achieved for the standard bigrams Fig.\ref{fig:microbenchmarks}).
\sys achieves a COST~\cite{mcsherryscalability} of 2 for all benchmarks.\str
% \kk{Do we want to talk about cost without mentioning it further?}
%   size grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
% The first number alone is interesting, as it captures \sys's output for the sequential script.
% As there is no parallelism involved, it highlights the initial (bare minimum) cost corresponding to using \sys.
% % this is the size of 
% This change in size does not necessarily translate to observed runtime, as most of the size increase comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
% FIFOs are simply pipes made explicit;
%   the performance characteristics, including in-kernel buffering and synchronization mechanisms, are identical between the two.

\heading{Discussion}
%
As discussed in (\S\ref{eval}, Parallelism) \sys does not have any control over the input script parallelism, and therefore it often achieves the maximum possible speedup, \ie without unnecessary process spawns and context switches, earlier than the purported parallelism---$16-32\times$ for a 64-core system. This is also due to the fact that \sys also adds its runtime primitives (\ttt{split} and \ttt{eager}), further increasing process spawns.\str
% \kk{Can we somehow turn this point around and make it seems good/fine. That here we report this parallelism, but in the real system, it tries to do a best effort guess of parallelism based on the size of the input script.}
% \sys achieves near-linear speedup on low-parallelism configurations.
% This is due to a high degree (deep trees) of task parallelism from \sys's multiple constructs---\eg \emph{aggregations} and \ttt{relay}s---evident in all plots that contain \pur stages.
%% \kk{Ben's point: task parallelism is mentioned here for the first time. Since \sys exposes data parallelism as task parallelism, we might want to mention that somewhere earlier in the paper}
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \emph{aggregation} nodes, and 14 \ttt{relay} nodes.
% In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
% significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
% Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available---as is the case, \eg  with Sort-sort for $32\times$ and $64\times$ which has 314 and 634 nodes, respectively.
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).
% \kk{Mention bigrams-opt result now that it is removed from the plot.}
\todo{No Eager} and \todo{Blocking Eager} depict configurations without \sys's \ttt{eager} optimization and without \ttt{split}. As described in \Cref{optimizer}, these eager implementations perform worse than \sys's optimized relay \ttt{eager} in all cases (except for a negligible difference in shortest-scripts). The benefits of \ttt{split} are also eveident for scripts that have \tpur or \tnpu commands, and for the rest it does not negatively affect performance.

>>>>

% To have any hope of parallelizing scripts, \sys first needs to understand the parallelization characteristics of the commands they contain.
\sys aims at parallelizing data-parallel commands, \ie ones that can process their input in parallel, encoding their characteristics by assigning them to
\emph{parallelizability classes}.
% \kk{eurosys: I have two small concerns about this sentence but I don't know how to fix them. First, there is a redundancy "parallelizing data-parallel". Second and most important, I would like this sentence to say that \sys exploits latent data parallelism in commands, but not that it parallelizes data parallel commands. The reason is that I would like to make sure that no reviewer thinks that all we do is parallelizing each command on its own (as some OSDI reviewer thought).}
%% This information is encoded in a command's \emph{parallelizability class}.
% Intuitively, a parallelizability 
% A class represents the level of synchronization required by copies of a command executing in parallel, 
%   capturing only characteristics that are important for its parallel execution. % , rather than the command's full observable behavior.
\sys leans towards having a few coarse classes rather than many detailed ones---among other reasons, to simplify their understanding and use by command developers.
% Make the connections that this is like a type, that doesn't capture everything, but is good-enough

@@

% As parallelizability classes form a hierarchy from most parallelizable to least parallelizable~\sx{cmd}, a command is classified by the class of its least parallelizable flag.
% For example, if a custom \ttt{trace-sort} command is invoked with flags \ttt{r}, \ttt{n}, and \ttt{k2} that are in \pur and a flag \ttt{d} in \sid that writes debugging output to a file, it ends up in \sid.

@@

\heading{Example Annotations}
The parallelizability properties of each command are specified in an annotation record that contains several components, each addressing one of the key concerns.
On the top level, each annotation record contains a set of clauses, each one identified by a predicate on the command arguments (Concern (iii)). 
Each clause contains an assignment for the command, indicating its class (Concern (i)), and the sequence of its inputs and outputs (Concern (ii)).
%  (a JSON object) that
% Annotations treat flag arguments differently from file arguments, by checking if they start with a \ttt{-}.
% \footnote{
%   Note that the single character \ttt{-} is not an option as it is most often used to designate that a command reads from its standard input.
%   \nv{
%     There are options of of the form --f "x; y", and BSD options do not take \ttt{-} (like \ttt{ps aux}).
%     I am also not sure how important this last sentence is, as it did not answer a question I had---I expected arguments to be anything: if soneome wants to put ``pizza'' the DSL should not disallow it.
%   }
% }
% The complete annotation grammar is shown in Appendix~\ref{alg}. \str

@@

\heading{Translation Pass}
%
%% \kk{It might be a good idea to change this subsection based on the new
%%   as-late-as-possible architecture where everything is wrapped and
%%   translated at the final moment. Ideally, this would become the main
%%   point of this subsection, as the translation algorithm is kind of
%%   redundant when having the distributable regions too.}
%
\sys's front-end performs a depth-first search on the AST of
the given shell program.  During this pass, it extends the
dataflow regions bottom-up, translating their
independent components to DFG nodes until a barrier
construct is reached. All AST subtrees not translatable to DFGs are kept as they are. The output of the translation pass is the
original AST where dataflow regions have been replaced with
DFGs. 
% and calls to \sys's runtime primitives~\sx{optimizer}.

@@

\heading{Highlights}
\sys accelerates ($1.92-61.1\times$) all scripts ($>50$, with $>200$ commands).
\kk{Add speedup separated: (i) transformations, (ii) runtime primitives.}
% Averages: [1.955223776839714, 3.478591807121036, 5.721328556478535, 8.604380509595906, 10.666921423438112, 13.246864076708423]
% No Eager Averages: [1.6327834370824543, 2.5481043878162444, 3.8623841529643173, 5.931851620007484, 7.466566647963421, 9.35752032439642]
% From the rest, only two see a slowdown, only because \sys's setup ($\sim1s$) is higher than their runtime ($\sim0.4s$).
\sys's optimized primitives offer benefits even against manually parallelized scripts. % (such as GNU sort, and GNU parallel).
% \sys's compilation pass is negligible (avg. $<1s$). %  for non trivial program execution.
In two large cases, tasks such as data download, extraction, and preprocessing---often implemented using shell scripts and outside the focus of specialized parallelization frameworks---take comparable time to the execution of the main computation, and can be significantly accelerated by \sys.
% In terms of manual effort,

@@

\heading{Results}
% TODO \nv{Need to change terminology in plots---files should be (attempted) parallelism, name benchmark, cat should be merge.}
Fig.~\ref{fig:microbenchmarks} presents \sys's speedup as a function of parallelism (2--64$\times$).
% (Bi-grams-opt is shown as the \ttt{opt} line in Fig.~\ref{fig:microbenchmarks}'s Bi-grams.)
Average speedups of the optimized \sys, \ie \ttt{eager} and \ttt{split} enabled, for \{2, 4, 8, 16, 32, 64\}$\times$-parallelism are \todo{1.97, 3.5, 5.78, 8.83, 10.96, and 13.47}, respectively.
% and the geometric means are 1.92, 3.39, 5.48, 7.94, 8.86, 8.51.}
% \kk{For the one-liners where split doesn't provide any benefit (or overhead) we don't show the PaSh w/o split line.}
%
% we experiment with a $200\times$ level of parallelism to see potential overheads
Tab.~\ref{tab:eval} shows that \sys's compilation time is negligible. % ($<5s$)
It also contains filter and bigrams-opt (not in Fig.~\ref{fig:microbenchmarks}). The first (filter) is used to evaluate \sys's impact on IO-intensive scripts (\sys achieves a 1.5-2.5$\times$ speedup). The second (bigrams-opt) is an optimized version of bigrams where we grouped 5 commands into one \tpur command thas has a more efficient $map$ and $aggregate$ implementation. \sys achieves up to x30 speedup for this script (significantly higher than what is achieved for the standard bigrams Fig.\ref{fig:microbenchmarks}).
\sys achieves a COST~\cite{mcsherryscalability} of 2 for all benchmarks.
%   size grows due to the resulting complexity of coordinating among parallel executions---multiple FIFOs per pipeline stage, merging and splitting, encoding of divide-and-conquer synchronization \etc
% The first number alone is interesting, as it captures \sys's output for the sequential script.
% As there is no parallelism involved, it highlights the initial (bare minimum) cost corresponding to using \sys.
% % this is the size of 
% This change in size does not necessarily translate to observed runtime, as most of the size increase comes from the setup and teardown of FIFOs responsible for interprocess (IPC) communication.
% FIFOs are simply pipes made explicit;
%   the performance characteristics, including in-kernel buffering and synchronization mechanisms, are identical between the two.

\heading{Discussion}
%
As discussed in (\S\ref{eval}, Parallelism) \sys does not have any control over the input script parallelism, and therefore it often achieves the maximum possible speedup, \ie without unnecessary process spawns and context switches, earlier than the purported parallelism---$16-32\times$ for a 64-core system. This is also due to the fact that \sys also adds its runtime primitives (\ttt{split} and \ttt{eager}), further increasing process spawns.
% \kk{eurosys: Can we somehow turn this point around and make it seems good/fine. That here we report this parallelism, but in the real system, it tries to do a best effort guess of parallelism based on the size of the input script.}
% \sys achieves near-linear speedup on low-parallelism configurations.
% This is due to a high degree (deep trees) of task parallelism from \sys's multiple constructs---\eg \emph{aggregations} and \ttt{relay}s---evident in all plots that contain \pur stages.
%% \kk{Ben's point: task parallelism is mentioned here for the first time. Since \sys exposes data parallelism as task parallelism, we might want to mention that somewhere earlier in the paper}
For example, Sort in $8\times$ spawns 37 nodes:
  8 \ttt{tr} nodes, 8 \ttt{sort} nodes, 7 \emph{aggregation} nodes, and 14 \ttt{relay} nodes.
% In effect, \sys hits the maximum number of parallel processes (and thus achieves optimal performance, \ie one without unnecessary process spawns and context switches) 
% significantly earlier than the purported parallelism---$16-32\times$ for a 64-core system.
% Asking for parallelism beyond that point degrades speedup, as overheads still increase but no further parallelism is available---as is the case, \eg  with Sort-sort for $32\times$ and $64\times$ which has 314 and 634 nodes, respectively.
% 
%  On the other hand opt-bigrams 64 has 255 nodes and the sys time stays
%   the same so I think that it might have to do with the fact that
%   constant costs rise a lot (since opt-bigrams 32 takes just 1
%   minute).
% \kk{Mention bigrams-opt result now that it is removed from the plot.}
\todo{No Eager} and \todo{Blocking Eager} depict configurations without \sys's \ttt{eager} optimization and without \ttt{split}. As described in \Cref{optimizer}, these eager implementations perform worse than \sys's optimized relay \ttt{eager} in all cases (except for a negligible difference in shortest-scripts). The benefits of \ttt{split} are also eveident for scripts that have \tpur or \tnpu commands, and for the rest it does not negatively affect performance.
